Crowd density estimation from Wi-Fi positioning data in the Amsterdam Arena

### 2016-06-24 PaperOnSmoothing/draft-31052016.tex
introduction updated. issues with positioning section started

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{verbatim}
\begin{document}

@@

In our approach, we combat the three above mentioned issues by exploiting the ubiquity of smart phones, as in ((4) (5), (6) from proposal?). Our living laboratory is the Amsterdam Arena(citation). First, we are estimating in real-time anonymously the positions of the visitors via the signal strengths of their phones to the  Wi-Fi routers in the venue. This technique has been presented in (cite Jan's slides), and is similar to trilateration. However, due to signal interference, irregular internet packet rates and systematic errors, this methodology is imprecise as well.  Thus, second, we apply the law of the large numbers to obtain an estimation of the crowd density, also in real-time. With the first step, we are dealing with issues 1) and 2) mentioned above. With the second step, which is the main contribution of this paper, we are handling issue 3) -  that is, with our method the precision of estimation actually increases as the crowd density increases.

@@

@@

\subsection{Issues with positioning (title to reconsider)}

>>>>

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{verbatim}
\usepackage{graphicx}
\begin{document}

@@

In our approach, we combat the three above mentioned issues by exploiting the ubiquity of smart phones, as in ((4) (5), (6) from proposal?). Our living laboratory is the Amsterdam Arena(citation). Or methodology consists of three steps.  First, we are using the Wi-Fi routers in the stadium to gather the strengths of the wireless antena signals from individuals smart phones.  Second, we are estimating anonymously the positions of the visitors from the signal strengths in real time. This technique has been presented in (cite Jan's slides), and is similar to trilateration, that is, the way the Global Positioning System (GPS) is implemented. However, due to signal interference and blockage in dense crowds, irregular internet packet rates and systematic errors leading to multiple local optima, the positioning itself does not suffice to estimate the crowd density. Thus, in our third step, we aggregate the results from the positioning and apply the law of the large numbers to obtain an estimation of the crowd density, also in real-time. With the first two steps, we are dealing with issues 1) and 2) mentioned above. With the third step, which is the main contribution of this paper, we are handling issue 3) -  that is, with our method the precision of estimation actually increases as the crowd density increases.

The rest of the paper is organized as follows.....*******

@@

This section will be written later, maybe with input from Jan.

@@

Sonja:Mention somewhere that in the documentary of the LP disaster people were raising their phones up to look for a way out. (which is good , less blockage of signals)
\subsection{Issues with positioning (title to reconsider)}

Under ideal circumstances, the positioning itself would suffice to estimate the crowd density per region of interest:  every few seconds we would only need to count the number of estimates that fall in the region of interest. However, we have observed several issues, which are not related to the mathematical methodology behind the positioning step, that prevent us to apply direct counting. 
\begin{enumerate}
	\item
	Due to signal interference, when the optimization of the estimation of location of a certain MAC address is performed, there can be multiple local optima that are equally good candidates. For example, consider Fig. \ref{fig:trilateration_problem}. In the center of every ring there is a Wi-Fi router, that has estimated signal strength to a certain MAC device with a certain error range. The error range is represented by the thickness of the ring. Then, there are two possible regions which are equally good candidates for the positioning of the MAC device, and those regions are the two darkest regions where all three rings overlap. Note that the problem can arise regardless of the thickness of the rings or the number of Wi-Fi routers. 	
	\begin{figure}[h!]
		\centering
		\includegraphics[width=40mm]{trilateration_problem.jpg}
		\caption{Trilateration can lead to multiple local optima}
		\label{fig:trilateration_problem}
	\end{figure}
	In fact, we have taken a static MAC device and plotted the estimations of its coordinates through time, where we have plotted only the estimations with a relatively small error  (Fig. \ref{fig:bimodal})
	\begin{figure}[h!]
		\centering
		\includegraphics[width=100mm]{bimodal.jpg}
		\caption{Trilateration can lead to multiple local optima (cut the lower part from the picture)}
		\label{fig:bimodal}
	\end{figure} 
	The resulting plot (for the x-coordinate) reveals a bimodal distribution of the estimation. This phenomenon can be explained by the multiple local optima situation.

\item Volatility of packet rates. When a MAC device is connected to the WI-Fi internet, it sends packets with a relatively stable rate. However, during concert hours, very few devices are actually connected to the internet, and most of our estimations of positions come from signals that the device sends in a "probing" mode. In this case the packet rate is quite volatile, ranging from a few seconds to a few minutes (some citation?).   
\end{enumerate}
